### 2023-06-11 docs/research/LaTeX/research.tex
Add 4 charts

\documentclass[12pt, a4paper]{article}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{url}
\usepackage{cite}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\graphicspath{{images/}}
\title{GDPR: Is non-compliance profitable for Meta Platforms, Inc.?}
\author{Stoyan L. Kostadinov}
\date{June 2023}
\begin{document}
\maketitle

\begin{center}
    \includegraphics[width=0.3\textwidth]{fontys-logo}
\end{center}

\begin{abstract}
    % TODO
\end{abstract}

\subsection*{Keywords}

GDPR, Facebook, Meta, profit, non-compliance, privacy, financial impact

\section*{Introduction}

In the recent years, the increasing usage of the Internet has lead to a rapid
expansion of data collection and processing\cite{khan2014big}. As concerns
regarding data privacy and protection grew, the European Union (EU) implemented
the General Data Protection Regulation (GDPR) in May
2018\cite{greengard2018weighing, historyGdpr}.

GDPR aimed to safeguard the privacy rights of individuals within the EU by
regulating the collection, storage, and transfer of personal
data\cite{historyGdpr}. Notably, data should be collected for specified,
explicit and legitimate purposes\cite{europeanParliamentGdprArticle5} and data
can be transferred outside the EU if general data processing principles are in
place \cite{europeanParliamentGdprArticle44}. Compliance with GDPR has become a
critical issue for big-tech companies, which often handle massive amounts of
user data.

One of those companies is \textit{Meta Platforms, Inc.} (formerly
\textit{Facebook, Inc.}), which has been fined numerous times for GDPR
violations\cite{mrevzar2023analysis}. This has raised questions about the impact
of GDPR on \textit{Meta's} performance.

The goal of this paper is to investigate the financial impact of GDPR
non-compliance on \textit{Meta Platforms, Inc.} and to determine whether it is
profitable for the company to continue violating the regulations.

\section*{Method}

All numerical data about \textit{Meta Platforms, Inc.} has been gathered from
the official \textit{Meta Investor Relations} website\cite{fbMetaFinancials} -
\textbf{Balance Sheets}, \textbf{Income Statements} and \textbf{Form 10-Q}.

GDPR fines imposed on \textit{Meta Platforms, Inc.} have been gathered from the
\textit{GDPR Enforcement Tracker}
website\cite{enforcementtrackerGDPREnforcement} and filtered to include fines
imposed on their family of applications (\textbf{\textit{Facebook}},
\textbf{\textit{WhatsApp}}, \textbf{\textit{Instagram}}) and
\textbf{\textit{Meta Platforms, Inc.}} itself.

\subsection*{Inclusion criteria}

\begin{itemize}
    \item Articles published in English
    \item Articles published between 2018 and 2023
    \item Articles that discuss GDPR
    \item Articles that discuss Meta Platforms, Inc.
    \item Data between 2018 and 2023
\end{itemize}

\subsection*{Exclusion criteria}

\begin{itemize}
    \item Articles published in languages other than English
    \item Articles published before 2018
    \item Articles that do not discuss GDPR
    \item Articles that do not discuss Meta Platforms, Inc.
    \item Articles that discuss about ethics
    \item Data before 2018
\end{itemize}

\subsection*{Ethical considerations}

This study is approved by Fontys University of Applied Sciences, Eindhoven, The
Netherlands. All data is collected from publicly available sources and no
personal data is collected. The author is not affiliated with Meta Platforms,
Inc. or any other company mentioned in this study.

\section*{Results}

\subsection*{Fines analysis}

%Explore the fines imposed on Meta Platforms by regulatory authorities due to
%GDPR violations. Examine the magnitude of the fines, both in absolute terms and
%in relation to the company's financial performance. Compare the fines to Meta
%Platforms' revenue, profits, or other financial indicators to assess their
%significance.

% TODO: cite

\begin{figure}[h]
    \centering
    \includegraphics[width=1.00\textwidth]{monetarial-amount-of-gdpr-fines}
    \caption{Monetarial amount of fines issued to Meta Platforms, Inc.(in
    millions of
    \$)\cite{gdprFine1,gdprFine2,gdprFine3,gdprFine4,gdprFine5,gdprFine6,gdprFine7,gdprFine8,gdprFine9}}
    \label{fig:amount-of-gdpr-fines}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=1.00\textwidth]{violations-by-article}
    \caption{Number of violations by GDPR
    article\cite{gdprFine1,gdprFine2,gdprFine3,gdprFine4,gdprFine5,gdprFine6,gdprFine7,gdprFine8,gdprFine9}}
    \label{fig:violations-by-article}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=1.00\textwidth]{net-income}
    \caption{Net Income of Meta Platforms, Inc.(in billions of
    \$)\cite{2023q1,2020q4,2020q3,2020q2,2020q1,2019q4,2019q3,2019q2,2019q1,2018q4,2018q3,2018q2}}
    \label{fig:net-income}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=1.00\textwidth]{revenue}
    \caption{Revenue of Meta Platforms, Inc.(in billions of
    \$)\cite{2023q1,2020q4,2020q3,2020q2,2020q1,2019q4,2019q3,2019q2,2019q1,2018q4,2018q3,2018q2}}
    \label{fig:revenue}
\end{figure}

\subsection*{Investor Confidence and Stock Performance}

% Assess the impact of GDPR non-compliance on Meta Platforms' stock performance.
% Analyze stock price movements, market reactions to GDPR-related news, and
% changes in investor sentiment. Explore how GDPR issues have influenced
% investor confidence and the company's market capitalization.

\subsection*{Advertiser Reactions}

% Study the response of advertisers to GDPR non-compliance by Meta Platforms.
% Examine whether there has been a shift in advertiser behavior, such as reduced
% advertising spend, contract cancellations, or demands for stricter data
% protection measures. Analyze the financial implications of these reactions on
% Meta Platforms' advertising revenue.

% Average revenue per person Trends in Our Monetization by Facebook User
% Geography page 46 - advertising

\subsection*{User Base and User Engagement}

% Investigate whether GDPR non-compliance has had any impact on Meta Platforms'
% user base and user engagement metrics. Analyze changes in user growth rates,
% active user counts, or user activity patterns following GDPR-related
% incidents. Assess the potential financial consequences of any decline in user
% metrics.

% Daily Active People Monthly Active People Daily Active Users (DAUs) facebook -
% per region Monthly Active Users (MAUs) facebook - per region

% https://d18rn0p25nwr6d.cloudfront.net/CIK-0001326801/5fb5d0ea-c2c7-46f0-a26a-656df1673fac.pdf
% page 14 revenue from europe page 42 41 37 and onwards

\section{Discussion}

% % Investigate the costs incurred by Meta Platforms in legal proceedings
% related to GDPR non-compliance. This includes expenses associated with hiring
% legal counsel, litigation, settlement agreements, and any ongoing legal
% obligations. Analyze the financial impact of these legal expenses on the
% company's profitability. Reputational Damage Assess the financial impact of
% reputational damage resulting from GDPR non-compliance. Investigate whether
% negative media coverage, public backlash, or loss of user trust have affected
% Meta Platforms' brand value or customer perception. Analyze any associated
% costs, such as brand recovery campaigns or decreased user acquisition rates.

% hard to assess

% https://www.law.com/corpcounsel/2021/11/24/as-regulators-hover-meta-expands-legal-department/

% https://www.law.com/corpcounsel/2022/11/10/meta-cuts-hit-legal-department-slamming-brakes-on-its-explosive-growth/?slreturn=20230510140111

% https://www.google.com/search?channel=fs&client=ubuntu&q=glasdoor+meta+lawyer+salary

\bibliographystyle{vancouver}
\bibliography{refs}
\end{document}